Viscous Transport in Eroding Porous Media

### 2019-10-09 docs/transport/erosion_transport.tex
worked on abstract

\begin{abstract} 
  We extend our previous work, [{\em Journal of Computational Physics},
  {\bf 375}:1--21, 2018], to simulate fluid-mechanical erosion of
  two-dimensional porous media with nearly-touching grains.  The method
  continues to use a boundary integral equation (BIE) method to solve
  the fluid equations, and a stable second-order time stepping method to
  erode the grains.  In our previous work, we required eroding grains
  that were sufficiently separated to guarantee that the trapezoid rule
  provides a sufficient amount of accuracy. We now remove this
  restriction by introducing a Barycentric quadrature rule whose error
  is bounded independent of the geometry.  The Barycentric quadrature
  rule is computationally expensive, so it is only used when necessary,
  and the spectrally accurate fast-multipole-accelerated trapezoid rule
  is used otherwise. We also simulate and analyze transport through
  eroded porous media.  By using the Barycentric quadrature rule, the
  analysis examines flow both in the fast and slow regions with
  high-order accuracy. Our results reveal several interesting  effects
  of erosion on the tortuosity, dispersion, and pore size distributions.
\end{abstract}

\begin{keyword}
  Porous media \sep Stokes flow \sep Erosion \sep Quadrature \sep
  Transport \sep Tortuosity \sep Anomalous dispersion \sep Pore size
  distributions
\end{keyword}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b1}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b80}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b160}\\
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b240}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b320}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b400}
\caption{\label{fig:Eroding50vel} Six evenly spaced snapshots of 50
eroding bodies.  The background velocity is a Hagen-Poiseuille flow with
a varying flow strength that corresponds to a constant pressure drop
across the channel. The color is the magnitude of the velocity in a
logarithmic scale. The flow velocity varies over several orders of
magnitude, the grains form irregular shapes with large aspect ratios,
and the geometry becomes channelized and anisotropic.}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b1}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b80}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b160}\\
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b240}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b320}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b400}
\caption{\label{fig:Eroding50vort} The same six snapshots as
Figure~\ref{fig:Eroding50vel}.  The color is the vorticity of the fluid.
Since the rate of erosion is equivalent to the magnitude of the
vorticity, erosion is fastest in the yellow and blue regions and slowest
in the green regions.}
\end{center}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pore Throat Size}
\label{sec:throats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
While transport in porous media depends on the porosity, it also depends
on the placement of the grains.  In particular, grain placement affects
velocity scales~\cite{ali-par-wei-bre2017}, correlation
structures~\cite{leb-ded-dav-bou2007}, contaminant
transport~\cite{knu-car2005},
channelization~\cite{sie-ili-pri-riv-gua2019,berhanu2012shape}, and pore
network models~\cite{bry-kin-mel1993, bry-mel-cad1993, bij-blu2006}. To
characterize the grain placement, we compute distributions of pore sizes
between neighboring grains.  To define neighboring grains, we form a
Delaunay triangulation using nodes placed at the center of each eroding
grain and at a collection of points around the boundary of the porous
media.  Then, we say that two grains are neighbors if their centers
share an edge of the triangulation~\cite{dea-qua-bir-jua2018}. The pores
of an eroded geometry are illustrated in
Figure~\ref{fig:Eroding100gap}(a). We do not consider pores between
eroding bodies and the solid wall $\Gamma$, so some of the grains near
the porous region boundary only have two neighbors. Having defined the
pore sizes, we plot its distribution in
Figure~\ref{fig:Eroding100gap}(b) and compare it with the Weibull
distribution, a distribution used by others to characterize pore
sizes~\cite{ioa-cha1993}.  In Section~\ref{sec:Eroding100}, we
investigate the effect of erosion on the pore size distribution.
\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a) The pore sizes between
neighboring grains. Two grains are said to be neighbors if they share an
edge of a Delanuay triangulation with nodes at the center of each
eroding grain. (b) The distribution of the pore sizes of an eroded
geometry. The black curve is the Weibull distribution with the same
first two moments as the data.}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r1h_shear}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d5h_shear}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.38\textwidth}
\includegraphics[width =\textwidth]{./figs/1b_0d4r0d1h_shear}
\caption{}
\end{subfigure}
\caption{\label{fig:NearWall} A single body eroding in a shearing Stokes
flow.  The color is the logarithm of the shear stress. Therefore,
erosion is fastest in the red regions (upper half) and slowest in the
blue regions (lower half).  The body is initialized at three different
distances from the lower wall: (a) $h$, (b) $h/2$, and (c) $h/10$.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r1h_vort}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d5h_vort}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d1h_vort}
\caption{}
\end{subfigure}
\caption{\label{fig:NearWall_vort} The vorticity of the fluid with a
single body eroding at time $t=0.1$. The initial distance from the body
to the solid wall are: (a) $h$, (b) $h/2$, and (c) $h/10$.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo[inline]{Bryan is here}
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points.  The background flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 8. Therefore, $U=1$ once all the bodies have
vanished.  Snapshots of the bodies at four equispaced times are shown in
Figure~\ref{fig:Eroding20vort}.  The color is the vorticity, whose
absolute value is proportional to the erosion rate on solid boundaries.
Initially, several of the eroding bodies are closer to the outer wall
than the $5h$ threshold required for the trapezoid rule to achieve
machine precision.  In particular, the distance between bodies 1, 6, 13,
and 15 and the outer wall is $1.3h$, $2.9h$, $2.8h$, and $1.3h$,
respectively, where $h$ is the arclength spacing of the outer wall
$\Gamma$.  In addition, the distance between several pairs of eroding
bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and 14 \& 18, is too small to
be accurately resolved with the trapezoid rule.  By using the
Barycentric quadrature rule, the interaction between these
nearly-touching bodies can be accurately resolved to the desired
accuracy, and erosion can be simulated until all the bodies have
vanished.

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} The erosion of 20 nearly touching
bodies, each discretized with $N_\iin=256$ points, in a Hagen-Poiseuille
flow.  The four snapshots are evenly spaced in time, and the color is
the fluid vorticity. In the fourth frame, bodies 4, 12, and 17 have
vanished completely, and body 19 has almost completely eroded.}
\end{center}
\end{figure}

Erosion causes the space between nearly touching bodies to quickly
expand, and flat faces develop along the regions of near contact.  This
qualitative behavior is seen in Figure~\ref{fig:Eroding20vort} between
bodies 3 \& 4, 15 \& 16, and was also observed in previous work.
However, by resolving the interaction between bodies that are much
closer together than in previous work, we observe that very little
erosion occurs between certain pairs of bodies, at least initially.  For
instance the opening between bodies 1 \& 6, 3 \& 9, and 5 \& 13 grow
much slower than the opening between bodies 15 \& 16.  A common feature
of the pores that grow slowly is that they are nearly perpendicular to
the main flow direction, resulting in a small erosion rate.

Following our previous work~\cite{qua-moo2018}, we begin by analyzing
the effect of erosion on the area fraction and the flow rate.  In
Figure~\ref{fig:Eroding20flowrate}(a), we plot the area fraction as a
function of normalized time.  The general trend of the area fraction
resembles our previous work~\cite{qua-moo2018} (Figure 10(a)), but with
a larger initial area fraction.  In
Figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\cite{qua-moo2018} (Figure 10(b)), except that the initial flow
rate is an order of magnitude smaller because of the larger initial area
fraction.  Starting around normalized time $0.2$,
Figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as a exponential law.  The line of
best fit is
\begin{align} 
  \log_{10} U \approx 4.32 \frac{t}{t_f} + c, 
\end{align}
$t_f$ is the time required for all the bodies to erode, and $c$ is a
fitting parameter. By construction $U = 1$ when $t=t_f$, so $c=-4,32$,
and the flow rate is approximated as
\begin{align}
  U \approx \exp\left(9.94 \left(\frac{t - t_f}{t_f} \right)\right),
\end{align}
which is the dashed line in Figure~\ref{fig:Eroding20flowrate}(b).

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/flow_rate20densen}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a) The area fraction of 20
nearly touching eroding bodies versus normalized time. (b) The flow
rate, $U$, for a fixed pressure drop across the channel versus
normalized time.  The flow rate is initially very small, but it
eventually increases as a power law (dashed line) towards the flow rate
$U=1$ that occurs once all the bodies have eroded.}
\end{figure}

In Figure~\ref{fig:Eroding20vort}, we observe that erosion creates a
network of channels from the inlet to outlet where the velocity and
vorticity, and therefore erosion rate, are much larger than other
regions.  These channels can be further visualized by considering the
streamlines.  In Figure~\ref{fig:Eroding20tracer}, we freeze the
geometry at the second snapshot from Figure~\ref{fig:Eroding20vort} and plot 200 streamlines that are initialized at $x=-1$ and equally
spaced in $y \in (-1,1)$.  The streamlines are shown at five different
times, and the final plot is a zoom in of the lower right quadrant of
the fifth time step, but with additional streamlines.  Since we use a
high-order quadrature rule and time stepping method, we resolve
streamlines that come very close to the eroded bodies.  There are three
clear regions where the streamlines are most concentrated, corresponding
to the regions of highest velocity.  Two of these regions are located
between the bodies and the solid walls at $y=\pm 1$, and the third cuts
through the porous region with the upper part of the channel formed by
bodies 1, 4, 6, and 8.  Since the flow is fastest in these regions, the
shearing is largest, and this causes the channels to continue to open
fastest as observed in Figure~\ref{fig:Eroding20vort}.

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150n}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from Figure~\ref{fig:Eroding20vort}. The streamlines are
initialized at $x=-1$ and are equally spaced in $y \in (-1,1)$. The
first five snapshots are evenly spaced in time.  The bottom right frame
is a zoom in of the fifth snapshot, but with additional streamlines.
Since we use the Barycentric quadrature rule and the fourth-order
Runge-Kutta method, some streamlines are very close to the grains
without penetrating them.}
\end{center}
\end{figure}

Next, we use the streamlines to compute the tortuosity and the anomalous
dispersion rates of eroding geometries.  To achieve statistical
convergence, we use $N_p = 1000$ streamlines~\cite{bel-sal-rin1992}. To
compute the tortuosity, we require the velocity at the inlet $x=-1$.
These normalized velocities are plotted in
Figure~\ref{fig:Eroding20tort}(a) for the eroded geometry at porosity
$\phi = 62.9\%$ (Figure~\ref{fig:Eroding20tort}(c)).  The velocities are
similar to the work of Matyka et al.~\cite{mat-kha-koz2008} (see Figure
4(a)), except that our cross-section, by construction, does not cut
through any of the grains.  Next, in Figure~\ref{fig:Eroding20tort}(b),
we plot the local tortuosity~\eqref{eqn:localTort} by calculating the
relative length of each streamline as it traverses the channel from
$x=-1$ to $x=1$.  The local tortuosity ranges from 1 to 1.27, meaning
that one of the streamlines is 27\% longer than it would have been if
the grains were absent.  The average streamline is 9.79\% longer or
equivalently the tortuosity of the geometry is $1.098$.  Again,
comparing the local tortuosity to Matyka~\cite{mat-kha-koz2008} (Figure
4(b)), the results are qualitatively similar. However, since our initial
cross-section does not cut through the grains, the local tortuosity does
not have any gaps.  Discontinuities in local tortuosity occur when
nearby streamlines diverge to circumvent a grain.  In
Figure~\ref{fig:Eroding20tort}(c), we plot pairs of streamlines
associated with the ten largest jumps in the local tortuosity, with each
pair of corresponding streamlines plotted in the same color.

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =0.97\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity of a porous geometry
that initially had 20 grains and has eroded to a porosity of 62.9\%.
(a) The $x$-component of the velocity at the inlet, $u_1(-1,y)$,
normalized by its maximum velocity of $2.98 \times 10^{-3}$. (b) The
local tortuosity $\tau(y)$ on the cross section $x = -1$.  The spatially
averaged tortuosity through the channel is $1.098$.  (c) The streamlines
resulting in the ten largest differences between neighboring
streamlines.  Neighboring streamlines have the same color.}
\end{figure}

>>>>

\begin{abstract} 
  Transport of viscous fluid through porous media is a direct
  consequence of the pore structure. Here we investigate transport
  through specific two-dimensional porous geometries---those that have
  become channelized by the process of fluid-mechanical erosion.  We
  consider passive tracers in the eroded geometries and analyze their to
  statistical moments to investigate the effect of erosion on the
  tortuosity and dispersion rates. We compare the computed
  porosity-tortuosity relationship with models available in the
  literature. Finally, we analyze the pore size distribution of an
  eroding geometry. The simulations are performed by combining a
  high-fidelity boundary integral equation solver for the Stokes
  equations, a second-order stable time stepping method to simulate
  erosion, a new numerical method to stably and accurately compute the
  eroded geometries and particle trajectories.

%  We extend our previous work, [{\em Journal of Computational Physics},
%  {\bf 375}:1--21, 2018], to simulate fluid-mechanical erosion of
%  two-dimensional porous media with nearly-touching grains.  The method
%  continues to use a boundary integral equation (BIE) method to solve
%  the fluid equations, and a stable second-order time stepping method to
%  erode the grains.  In our previous work, we required eroding grains
%  that were sufficiently separated to guarantee that the trapezoid rule
%  provides a sufficient amount of accuracy. We now remove this
%  restriction by introducing a Barycentric quadrature rule whose error
%  is bounded independent of the geometry.  The Barycentric quadrature
%  rule is computationally expensive, so it is only used when necessary,
%  and the spectrally accurate fast-multipole-accelerated trapezoid rule
%  is used otherwise. We also simulate and analyze transport through
%  eroded porous media.  By using the Barycentric quadrature rule, the
%  analysis examines flow both in the fast and slow regions with
%  high-order accuracy. Our results reveal several interesting  effects
%  of erosion on the tortuosity, dispersion, and pore size distributions.
\end{abstract}

\begin{keyword}
  Porous media \sep Stokes flow \sep Erosion \sep Transport \sep
  Tortuosity \sep Anomalous dispersion \sep Pore size distributions \sep
  Quadrature
\end{keyword}

@@

\begin{figure}[htp]
\begin{center}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b1}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b80}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b160}\\
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b240}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b320}
\includegraphics[height = 0.31 \textwidth]{./figs/vel_log_50b400}
\caption{\label{fig:Eroding50vel} Six evenly spaced snapshots of 50
eroding bodies.  The background velocity is a Hagen-Poiseuille flow with
a varying flow strength that corresponds to a constant pressure drop
across the channel. The color is the magnitude of the velocity in a
logarithmic scale. The flow velocity varies over several orders of
magnitude, the grains form irregular shapes with large aspect ratios,
and the geometry becomes channelized and anisotropic.}
\end{center}
\end{figure}

@@

\begin{figure}[htp]
\begin{center}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b1}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b80}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b160}\\
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b240}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b320}
\includegraphics[height = 0.3 \textwidth]{./figs/vort_50b400}
\caption{\label{fig:Eroding50vort} The same six snapshots as
Figure~\ref{fig:Eroding50vel}.  The color is the vorticity of the fluid.
Since the rate of erosion is equivalent to the magnitude of the
vorticity, erosion is fastest in the yellow and blue regions and slowest
in the green regions.}
\end{center}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pore Throat Size}
\label{sec:throats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
While transport in porous media depends on the porosity, it also depends
on the placement of the grains.  In particular, grain placement affects
velocity scales~\cite{ali-par-wei-bre2017}, correlation
structures~\cite{leb-ded-dav-bou2007}, contaminant
transport~\cite{knu-car2005},
channelization~\cite{sie-ili-pri-riv-gua2019,berhanu2012shape}, and pore
network models~\cite{bry-kin-mel1993, bry-mel-cad1993, bij-blu2006}. To
characterize the grain placement, we compute distributions of pore sizes
between neighboring grains.  To define neighboring grains, we form a
Delaunay triangulation using nodes placed at the center of each eroding
grain and at a collection of points around the boundary of the porous
media.  Then, we say that two grains are neighbors if their centers
share an edge of the triangulation~\cite{dea-qua-bir-jua2018}. The pores
of an eroded geometry are illustrated in
Figure~\ref{fig:Eroding100gap}(a). We do not consider pores between
eroding bodies and the solid wall $\Gamma$, so some of the grains near
the porous region boundary only have two neighbors. Having defined the
pore sizes, we plot its distribution in
Figure~\ref{fig:Eroding100gap}(b) and compare it with the Weibull
distribution, a distribution used by others to characterize pore
sizes~\cite{ioa-cha1993}.  In Section~\ref{sec:Eroding100}, we
investigate the effect of erosion on the pore size distribution.
\begin{figure}[htp]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a) The pore sizes between
neighboring grains. Two grains are said to be neighbors if they share an
edge of a Delanuay triangulation with nodes at the center of each
eroding grain. (b) The distribution of the pore sizes of an eroded
geometry. The black curve is the Weibull distribution with the same
first two moments as the data.}
\end{figure}

@@

\begin{figure}[htp]
\begin{center}
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[height = 0.74\textwidth]{./figs/1b_0d4r1h_shear}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[height = 0.74\textwidth]{./figs/1b_0d4r0d5h_shear}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[height = 0.74\textwidth]{./figs/1b_0d4r0d1h_shear}
\caption{}
\end{subfigure}
\end{center}
\caption{\label{fig:NearWall} A single body eroding in a shearing Stokes
flow.  The color is the logarithm of the shear stress. Therefore,
erosion is fastest in the red regions (upper half) and slowest in the
blue regions (lower half).  The body is initialized at three different
distances from the lower wall: (a) $h$, (b) $h/2$, and (c) $h/10$.}
\end{figure}
\begin{figure}[htp]
\begin{center}
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[height = 0.53\textwidth]{./figs/1b_0d4r1h_vort}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[height = 0.53\textwidth]{./figs/1b_0d4r0d5h_vort}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[height = 0.53\textwidth]{./figs/1b_0d4r0d1h_vort}
\caption{}
\end{subfigure}
\caption{\label{fig:NearWall_vort} The vorticity of the fluid with a
single body eroding at time $t=0.1$. The initial distance from the body
to the solid wall are: (a) $h$, (b) $h/2$, and (c) $h/10$.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points.  The background flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the average pressure drop from
$x=-2$ to $x=2$ is held fixed at 8. Therefore, $U=1$ once all the bodies
have vanished.  Snapshots of the bodies at four equispaced times are
shown in Figure~\ref{fig:Eroding20vort}.  The color is the vorticity,
whose absolute value is proportional to the erosion rate on solid
boundaries.  Initially, several of the eroding bodies are closer to the
outer wall than the $5h$ threshold required for the trapezoid rule to
achieve machine precision.  In particular, the distance between bodies
1, 6, 13, and 15 and the outer wall is $1.3h$, $2.9h$, $2.8h$, and
$1.3h$, respectively, where $h$ is the arclength spacing of the outer
wall $\Gamma$.  In addition, the distance between several pairs of
eroding bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and 14 \& 18, is too
small to be accurately resolved with the trapezoid rule.  By using the
Barycentric quadrature rule, the interaction between these
nearly-touching bodies is accurately resolved to the desired accuracy,
and erosion can be simulated until all the bodies have vanished.

\begin{figure}[htp]
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} The erosion of 20 nearly touching
bodies, each discretized with $N_\iin=256$ points, in a Hagen-Poiseuille
flow.  The four snapshots are evenly spaced in time, and the color is
the fluid vorticity. In the fourth frame, bodies 4, 12, and 17 have
vanished completely, and body 19 has almost completely eroded.}
\end{center}
\end{figure}

Erosion causes the space between nearly touching bodies to quickly
expand, and flat faces develop along the regions of near contact.  This
qualitative behavior is seen in Figure~\ref{fig:Eroding20vort} between
bodies 3 \& 4, 15 \& 16, and was also observed in previous
work~\cite{qua-moo2018}. However, by resolving the interaction between
bodies that are much closer together than in previous work, we observe
that very little erosion occurs between certain pairs of bodies, at
least initially.  For instance the opening between bodies 1 \& 6, 3 \&
9, and 5 \& 13 grow much slower than the opening between bodies 15 \&
16.  A common feature of the pores that grow slowly is that they are
nearly perpendicular to the main flow direction, resulting in a small
erosion rate.

\begin{figure}[htp]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/flow_rate20densen}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a) The area fraction of 20
nearly touching eroding bodies versus normalized time. (b) The flow
rate, $U$, for a fixed pressure drop across the channel versus
normalized time.  The flow rate is initially very small, but it
eventually increases as an exponential law  (dashed line) towards the
flow rate $U=1$ that occurs once all the bodies have eroded.}
\end{figure}

We next analyze the effect of erosion on the area fraction and the flow
rate.  In Figure~\ref{fig:Eroding20flowrate}(a), we plot the area
fraction as a function of normalized time.  The general trend of the
area fraction resembles our previous work~\cite{qua-moo2018} (Figure
10(a)), but with a larger initial area fraction.  In
Figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\cite{qua-moo2018} (Figure 10(b)), except that the initial flow
rate is an order of magnitude smaller because of the larger initial area
fraction.  Starting around normalized time $0.2$,
Figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as an exponential law.  The line of
best fit is
\begin{align} 
  \log_{10} U \approx 4.32 \frac{t}{t_f} + c, 
\end{align}
where $t_f$ is the time required for all the bodies to erode, and $c$ is
a fitting parameter. By construction $U = 1$ when $t=t_f$, so $c=-4.32$,
and the flow rate is approximated as
\begin{align}
  U \approx \exp\left(9.94 \left(\frac{t - t_f}{t_f} \right)\right),
\end{align}
which is the dashed line in Figure~\ref{fig:Eroding20flowrate}(b).

\begin{figure}[htp]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150n}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from Figure~\ref{fig:Eroding20vort}. The streamlines are
initialized at $x=-1$ and are equally spaced in $y \in (-1,1)$. The
first five snapshots are evenly spaced in time.  The bottom right frame
is a zoom in of the fifth snapshot, but with additional streamlines.
Since we use the Barycentric quadrature rule and a fourth-order
Runge-Kutta method, some streamlines are very close to the grains
without penetrating them.}
\end{center}
\end{figure}

In Figure~\ref{fig:Eroding20vort}, we observe that erosion creates a
network of channels from the inlet to outlet where the velocity and
vorticity, and therefore erosion rate, are much larger than other
regions.  These channels can be further visualized  with the
streamlines.  In Figure~\ref{fig:Eroding20tracer}, we freeze the
geometry at the second snapshot from Figure~\ref{fig:Eroding20vort} and
plot 200 streamlines that are initialized at $x=-1$ and equally spaced
in $y \in (-1,1)$.  The streamlines are shown at five different times,
and the final plot is a zoom in of the lower right quadrant of the fifth
time step, but with additional streamlines.  Since we use a high-order
quadrature rule and time stepping method, we resolve streamlines that
come very close to the eroded bodies.  There are three clear regions
where the streamlines are most concentrated, corresponding to the
regions of highest velocity.  Two of these regions are located between
the bodies and the solid walls at $y=\pm 1$, and the third cuts through
the porous region with the upper part of the channel formed by bodies 1,
4, 6, and 8.  Since the flow is fastest in these regions, the shearing
is largest, and this causes the channels to continue to open fastest as
observed in Figure~\ref{fig:Eroding20vort}.

Next, we use the $N_p = 1000$~\cite{bel-sal-rin1992} streamlines to
compute the tortuosity of the eroding geometry.  To compute the
tortuosity, we require the velocity at the inlet $x=-1$.  These
normalized velocities are plotted in Figure~\ref{fig:Eroding20tort}(a)
for the eroded geometry at porosity $\phi = 62.9\%$
(Figure~\ref{fig:Eroding20tort}(c)).  The velocities are similar to the
work of Matyka et al.~\cite{mat-kha-koz2008} (see Figure 4(a)), except
that our cross-section, by construction, does not cut through any of the
grains.  Next, in Figure~\ref{fig:Eroding20tort}(b), we plot the local
tortuosity~\eqref{eqn:localTort} by calculating the relative length of
each streamline as it traverses the channel from $x=-1$ to $x=1$.  The
local tortuosity ranges from 1 to 1.27, meaning that one of the
streamlines is 27\% longer than it would have been if the grains were
absent.  The average streamline is 9.79\% longer or equivalently the
tortuosity of the geometry is $1.098$.  Again, comparing the local
tortuosity to Matyka~\cite{mat-kha-koz2008} (Figure 4(b)), the results
are qualitatively similar. However, since our initial cross-section does
not cut through the grains, the local tortuosity does not have any gaps.
Discontinuities in local tortuosity occur when nearby streamlines
diverge to circumvent a grain.  In Figure~\ref{fig:Eroding20tort}(c), we
plot pairs of streamlines associated with the ten largest jumps in the
local tortuosity, with each pair of corresponding streamlines plotted in
the same color.

\begin{figure}[htp]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =0.97\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity of a porous geometry
that initially had 20 grains and has eroded to a porosity of 62.9\%.
(a) The $x$-component of the velocity at the inlet, $u_1(-1,y)$,
normalized by its maximum velocity of $2.98 \times 10^{-3}$. (b) The
local tortuosity $\tau(y)$ on the cross section $x = -1$.  The spatially
averaged tortuosity through the channel is $1.098$.  (c) The streamlines
resulting in the ten largest differences of local tortuosity between
neighboring streamlines.  Neighboring streamlines have the same color.}
\end{figure}
